### 2024-06-28 paper/tex/04-algorithm.tex
edit

\begin{figure}
  \begin{multicols}{2}
      Negative Algorithmic Variables\\
      $[[α̂⁻]]$, $[[β̂⁻]]$, $[[γ̂⁻]]$, \dots\\
      \columnbreak
      Positive Algorithmic Variables\\
      $[[α̂⁺]]$, $[[β̂⁺]]$, $[[γ̂⁺]]$, \dots\\
  \end{multicols}
  \hfill\\
  \begin{multicols}{2}
      \ottuNShort
      \columnbreak
      \ottuPShort
  \end{multicols}
  \hfill\\
  Algorithmic Type Context\\
   $[[Ξ]] \Coloneqq \{[[α1̂±]], \dots, [[αn̂±]]\}$ where $[[α1̂±]], \dots, [[αn̂±]]$ are pairwise distinct \\
  \hfill\\
  Instantiation Context\\
   $[[Θ]] \Coloneqq \{[[ α1̂±[Γ1] ]], \dots, [[ αn̂±[Γn] ]]\}$ where $[[α1̂±]], \dots, [[αn̂±]]$ are pairwise distinct \\
  \caption{Algorithmic Syntax}
  \label{fig:algo-syntax}
\end{figure}

@@

\begin{figure}
\begin{multicols}{2}
  $\vcenter{\hbox{\vdots}}$\\
  \ottusedrule{\ottdruleWFATPUVarLabeled{}}
  \columnbreak\\
  $\vcenter{\hbox{\vdots}}$\\
  \ottusedrule{\ottdruleWFATNUVarLabeled{}}
\end{multicols}
\caption{Well-formedness of Algorithmic Types}
\label{fig:algo-wf}
\end{figure}

@@

\begin{figure}
\begin{multicols}{2}
  $\vcenter{\hbox{\vdots}}$\\
  \ottusedrule{\ottdruleNrmPUVar{}}
  \columnbreak\\
  $\vcenter{\hbox{\vdots}}$\\
  \ottusedrule{\ottdruleNrmNUVar{}}
\end{multicols}
\caption{Normalization of Algorithmic Types}
\label{fig:algo-nf}
\end{figure}

@@

  \begin{figure}[h]
    \begin{multicols}{2}
      \ottgrammartabular { 
      \ottscE 
      \ottinterrule
      \ottruleheadOneLine
        {[[SC]]}{\Coloneqq} {\ottcom{Constraint Set}}
        {\{[[scE1]], \dots, [[scEn]]\}}\ottprodnewline
      }

    \columnbreak

      \ottgrammartabular { 
      \ottucE 
      \ottprodnewline
      \ottinterrule
      \ottruleheadOneLine
        {[[UC]]}{\Coloneqq} {\ottcom{Unification Constraint Set}}
        {\{[[ucE1]], \dots, [[ucEn]]\}}\ottprodnewline
      }
    \end{multicols}

    \label{fig:syntax-e-sc}
    \caption{Constraint Entries and Sets}
  \end{figure}

@@

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    [>={Stealth[scale=2]},node distance=2.4cm,every node/.style={draw,rectangle},every text node part/.style={align=center}]

@@

In the remainder of this section, we will delve into each of these procedures in
detail, following the top-down order of the dependency graph. First, 
we present the subtyping algorithm itself.

@@

\begin{figure}[h]
  \hfill\\
  \begin{multicols}{2}
    \ottdefnANsubLabeled{}
    \columnbreak\\
    \ottdefnAPsupLabeled{}
  \end{multicols}
  \caption{Subtyping Algorithm}
  \label{fig:alg-subtyping}
\end{figure}

@@

\begin{figure}[h]
  \hfill
  \begin{multicols}{2}
  \ottdefnUNUnifLabeled{}
  \columnbreak\\
  \ottdefnUPUnifLabeled{}
  \end{multicols}
  \caption{Unification Algorithm}
  \label{fig:unification}
\end{figure}

@@

\begin{figure}[h]
  \ottdefnSCMELabeled\\
  \caption{Merge of Matching Constraint Entries}
  \label{fig:merge-entries}
\end{figure}

@@

\begin{figure}[h]
  Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$.\\
  Then $[[Θ ⊢ SC1 & SC2 = SC]]$
  defines a set of constraints $[[SC]]$ such that $[[scE]] \in [[SC]]$ iff either:
  \begin{itemize}
    \item $[[scE]] \in [[SC1]]$ and there is no matching $[[scE']] \in [[SC2]]$; or
    \item $[[scE]] \in [[SC2]]$ and there is no matching $[[scE']] \in [[SC1]]$; or
    \item $[[Θ(α̂±) ⊢ scE1 & scE2 = scE]]$ for some $[[scE1]] \in [[SC1]]$ and $[[scE2]] \in [[SC2]]$
      such that $[[scE1]]$ and $[[scE2]]$ both restrict variable $[[α̂±]]$. 
  \end{itemize}

@@

\begin{figure}[h]
  \begin{multicols}{2}
    \ottdefnLUBUpLabeled{}
    \columnbreak\\
    \ottdefnLUBNsubLabeled{}
  \end{multicols}
  \caption{Type Upgrade and Leas Upper Bound Algorithms}
  \label{fig:type-upgrade}
\end{figure}

@@

\begin{figure}[h]
    \ottdefnAUAUPLabeled{}
    \hfill\\
    \hfill\\
    \ottdefnAUAUNLabeled{}
    \caption{Anti-Unification Algorithm}
    \label{fig:anti-unification}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    [>={Stealth[scale=2]},node distance=2.2cm,every node/.style={draw,rectangle},every text node part/.style={align=center}]
    % Define nodes
    \node[] (1) {Positive Inference\\$[[Γ; Φ ⊨ v : iP]]$};
    \node[right=1.4cm of 1] (2) {Negative Inference\\$[[Γ; Φ ⊨ c : iN]]$};
    \node[right=1.2cm of 2] (3) {Application Inference\\$[[Γ ; Φ ; Θ1 ⊨ uN ● args ⇒> uM ⫤ Θ2 ; SC]]$};
    \node[below=0.6cm of 2] (7) {Minimal Instantiation\\$[[Γ ⊢ uP SC minby uσ]]$\\(\cref{sec:constraint-singularity})};
    \node[below=0.5cm of 7] (8) {Constraint Singularity\\$[[SC singular with uσ]]$\\(\cref{sec:constraint-singularity})};
    \node[right=0.3cm of 7] (4) {Constraint Merge\\$[[Θ ⊢ SC1 & SC2 = SC3]]$\\(\cref{sec:constraint-merge})};
    \node[left=0.3cm of 7] (6) {Negative Subtyping\\$[[Γ ; Θ ⊨ uN1 ≤ iN2 ⫤ SC]]$\\(\cref{sec:subtyping-algorithm})};
    \node[below=0.9cm of 6] (5) {Positive Subtyping\\$[[Γ ; Θ ⊨ uP1 ≥ iP2 ⫤ SC]]$\\(\cref{sec:subtyping-algorithm})};
    \draw[->] ([yshift=-.2cm]2.west) to ([yshift=-.2cm]1.east);
    \draw[->] ([yshift=+.2cm]1.east) to ([yshift=+.2cm]2.west);
    \draw[->, bend right=20] (3) to (1);
    \draw[->] (2) to (3);
    \draw[->] (2) to (4);
    \draw[->] (2) to (7);
    \draw[->] (3) to (4);
    \draw[->] (7) to (8);
    \draw[->] (1.south west) |- (5);
    \draw[->] (3) |- (5.south east);
    \draw[->] (2) to (6);
  \end{tikzpicture}  
  \caption{Dependency graph of the typing algorithm}
  \label{fig:alg-typing-graph}
\end{figure}

@@

\begin{figure}
  \ottdefnATPInfLabeled{}
  \hfill \\
  \ottdefnATNInfLabeled{}
  \hfill \\
  \ottdefnATSpinInfLabeled{}
  \caption{Algorithmic Type Inference}
  \label{fig:type-inference}
\end{figure}

@@

\begin{figure}[h]
  \hfill\\
  \ottdefnMININSTLabeled
  \caption{Minimal Instantiation}
  \label{fig:minimal-instantiation}
\end{figure}

@@

\begin{figure}
  \begin{minipage}{0.49\textwidth}
    `$[[SC singular with uσ]]$' means:
    \begin{itemize}[leftmargin=*]
      \item[$+$] for any \emph{positive} constraint entry $[[scE ∊ SC]]$ restricting 
        a variable $[[β̂⁺]]$, there exists $[[iP]]$ such that $[[scE singular with iP]]$
        (as defined in \cref{fig:constraint-entry-singularity}),
        and $[[ [uσ]β̂⁺ = iP ]]$; and
      \item[$-$] the symmetric property holds for all \emph{negative} $[[scE ∊ SC]]$; and
      \item[$\notin$] for any $[[α̂± ∉ dom(SC)]]$, $[[ [uσ]α̂± = α̂± ]]$.
    \end{itemize}
    \caption{Singular Constraint}
    \label{fig:constraint-singularity}
  \end{minipage}
  \begin{minipage}{0.48\textwidth}
    \ottdefnSINGLabeled
    \caption{Singular Constraint Entry}
    \label{fig:constraint-entry-singularity}
  \end{minipage}
\end{figure}

>>>>

\begin{figure}[t]

  \begin{minipage}[t]{0.49\textwidth}
      Negative Algorithmic Variables\\
      $[[α̂⁻]]$, $[[β̂⁻]]$, $[[γ̂⁻]]$, \dots\\
  \end{minipage}%
  \begin{minipage}[t]{0.49\textwidth}
      Positive Algorithmic Variables\\
      $[[α̂⁺]]$, $[[β̂⁺]]$, $[[γ̂⁺]]$, \dots\\
  \end{minipage}

  \hfill\\
  \begin{minipage}[t]{0.49\textwidth}
      \ottuNShort
  \end{minipage}
  \begin{minipage}[t]{0.49\textwidth}
      \ottuPShort
  \end{minipage}
  \hfill\\
  Algorithmic Type Context\\
   $[[Ξ]] \Coloneqq \{[[α1̂±]], \dots, [[αn̂±]]\}$ where $[[α1̂±]], \dots, [[αn̂±]]$ are pairwise distinct \\
  \hfill\\
  Instantiation Context\\
   $[[Θ]] \Coloneqq \{[[ α1̂±[Γ1] ]], \dots, [[ αn̂±[Γn] ]]\}$ where $[[α1̂±]], \dots, [[αn̂±]]$ are pairwise distinct \\
  \caption{Algorithmic Syntax}
  \label{fig:algo-syntax}
\end{figure}

@@

\begin{figure}[t]
  \begin{minipage}{0.49\textwidth}
  $$\vcenter{\hbox{\vdots}}$$
  \ottusedrule{\ottdruleWFATPUVarLabeled{}}
  \end{minipage}
  \begin{minipage}{0.49\textwidth}
  $$\vcenter{\hbox{\vdots}}$$
  \ottusedrule{\ottdruleWFATNUVarLabeled{}}
  \end{minipage}
\caption{Well-formedness of Algorithmic Types extends Declarative Well-formedness}
\label{fig:algo-wf}
\end{figure}

@@

\begin{figure}[t]
\begin{minipage}{0.49\textwidth}
  $$\vcenter{\hbox{\vdots}}$$
  \ottusedrule{\ottdruleNrmPUVar{}}
\end{minipage}
\hfill
\begin{minipage}{0.49\textwidth}
  $$\vcenter{\hbox{\vdots}}$$
  \ottusedrule{\ottdruleNrmNUVar{}}
\end{minipage}
\caption{Normalization of Algorithmic Types extends Declarative Normalization}
\label{fig:algo-nf}
\end{figure}

@@

  \begin{figure}[t]
    \begin{minipage}{0.49\textwidth}
      \ottgrammartabular { 
      \ottscE 
      \ottinterrule
      \ottruleheadOneLine
        {[[SC]]}{\Coloneqq} {\ottcom{Constraint Set}}
        {\{[[scE1]], \dots, [[scEn]]\}}\ottprodnewline
      }
    \end{minipage}
    \begin{minipage}{0.49\textwidth}
      \ottgrammartabular { 
      \ottucE 
      \ottprodnewline
      \ottinterrule
      \ottruleheadOneLine
        {[[UC]]}{\Coloneqq} {\ottcom{Unification Constraint Set}}
        {\{[[ucE1]], \dots, [[ucEn]]\}}\ottprodnewline
      }
    \end{minipage}
    \label{fig:syntax-e-sc}
    \caption{Constraint Entries and Sets}
  \end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{tikzpicture}
    [>={Stealth[scale=2]},node distance=2.4cm,every node/.style={draw,rectangle},every text node part/.style={align=center}]

@@

In the remainder of this section, we will discuss each of these procedures in
detail, following the top-down order of the dependency graph. First, we present
the subtyping algorithm itself.

@@

\begin{figure}[t]
  \hfill\\
  \begin{multicols}{2}
    \ottdefnANsubLabeled{}
    \columnbreak\\
    \ottdefnAPsupLabeled{}
  \end{multicols}
  \caption{Subtyping Algorithm}
  \label{fig:alg-subtyping}
\end{figure}

@@

\begin{figure}[t]
  \hfill
  \begin{multicols}{2}
  \ottdefnUNUnifLabeled{}
  \columnbreak\\
  \ottdefnUPUnifLabeled{}
  \end{multicols}
  \caption{Unification Algorithm}
  \label{fig:unification}
\end{figure}

@@

\begin{figure}[t]
  \ottdefnSCMELabeled\\
  \caption{Merge of Matching Constraint Entries}
  \label{fig:merge-entries}
\end{figure}

@@

\begin{figure}[t]
  Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$.\\
  Then $[[Θ ⊢ SC1 & SC2 = SC]]$
  defines a set of constraints $[[SC]]$ such that $[[scE]] \in [[SC]]$ iff either:
  \begin{itemize}
    \item $[[scE]] \in [[SC1]]$ and there is no matching $[[scE']] \in [[SC2]]$; or
    \item $[[scE]] \in [[SC2]]$ and there is no matching $[[scE']] \in [[SC1]]$; or
    \item $[[Θ(α̂±) ⊢ scE1 & scE2 = scE]]$ for some $[[scE1]] \in [[SC1]]$ and $[[scE2]] \in [[SC2]]$
      such that $[[scE1]]$ and $[[scE2]]$ both restrict variable $[[α̂±]]$. 
  \end{itemize}

@@

\begin{figure}[t]
  \begin{multicols}{2}
    \ottdefnLUBUpLabeled{}
    \columnbreak\\
    \ottdefnLUBNsubLabeled{}
  \end{multicols}
  \caption{Type Upgrade and Leas Upper Bound Algorithms}
  \label{fig:type-upgrade}
\end{figure}

@@

\begin{figure}[t]
    \ottdefnAUAUPLabeled{}
    \hfill\\
    \hfill\\
    \ottdefnAUAUNLabeled{}
    \caption{Anti-Unification Algorithm}
    \label{fig:anti-unification}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{tikzpicture}
    [>={Stealth[scale=2]},node distance=2.2cm,every node/.style={draw,rectangle},every text node part/.style={align=center}]
    % Define nodes
    \node[] (1) {Positive Inference\\$[[Γ; Φ ⊨ v : iP]]$};
    \node[right=1.4cm of 1] (2) {Negative Inference\\$[[Γ; Φ ⊨ c : iN]]$};
    \node[right=1.2cm of 2] (3) {Application Inference\\$[[Γ ; Φ ; Θ1 ⊨ uN ● args ⇒> uM ⫤ Θ2 ; SC]]$};
    \node[below=0.6cm of 2] (7) {Minimal Instantiation\\$[[Γ ⊢ uP SC minby uσ]]$\\(\cref{sec:constraint-singularity})};
    \node[below=0.5cm of 7] (8) {Constraint Singularity\\$[[SC singular with uσ]]$\\(\cref{sec:constraint-singularity})};
    \node[right=0.3cm of 7] (4) {Constraint Merge\\$[[Θ ⊢ SC1 & SC2 = SC3]]$\\(\cref{sec:constraint-merge})};
    \node[left=0.3cm of 7] (6) {Negative Subtyping\\$[[Γ ; Θ ⊨ uN1 ≤ iN2 ⫤ SC]]$\\(\cref{sec:subtyping-algorithm})};
    \node[below=0.9cm of 6] (5) {Positive Subtyping\\$[[Γ ; Θ ⊨ uP1 ≥ iP2 ⫤ SC]]$\\(\cref{sec:subtyping-algorithm})};
    \draw[->] ([yshift=-.2cm]2.west) to ([yshift=-.2cm]1.east);
    \draw[->] ([yshift=+.2cm]1.east) to ([yshift=+.2cm]2.west);
    \draw[->, bend right=20] (3) to (1);
    \draw[->] (2) to (3);
    \draw[->] (2) to (4);
    \draw[->] (2) to (7);
    \draw[->] (3) to (4);
    \draw[->] (7) to (8);
    \draw[->] (1.south west) |- (5);
    \draw[->] (3) |- (5.south east);
    \draw[->] (2) to (6);
  \end{tikzpicture}  
  \caption{Dependency graph of the typing algorithm}
  \label{fig:alg-typing-graph}
\end{figure}

@@

\begin{figure}[t]
  \ottdefnATPInfLabeled{}
  \hfill \\
  \ottdefnATNInfLabeled{}
  \hfill \\
  \ottdefnATSpinInfLabeled{}
  \caption{Algorithmic Type Inference}
  \label{fig:type-inference}
\end{figure}

@@

\begin{figure}[t]
  \hfill\\
  \ottdefnMININSTLabeled
  \caption{Minimal Instantiation}
  \label{fig:minimal-instantiation}
\end{figure}

@@

\begin{figure}[t]
  \begin{minipage}{0.49\textwidth}
    `$[[SC singular with uσ]]$' means:
    \begin{itemize}[leftmargin=*]
      \item[$+$] for any \emph{positive} constraint entry $[[scE ∊ SC]]$ restricting 
        a variable $[[β̂⁺]]$, there exists $[[iP]]$ such that $[[scE singular with iP]]$
        (as defined in \cref{fig:constraint-entry-singularity}),
        and $[[ [uσ]β̂⁺ = iP ]]$; and
      \item[$-$] the symmetric property holds for all \emph{negative} $[[scE ∊ SC]]$; and
      \item[$\notin$] for any $[[α̂± ∉ dom(SC)]]$, $[[ [uσ]α̂± = α̂± ]]$.
    \end{itemize}
    \caption{Singular Constraint}
    \label{fig:constraint-singularity}
  \end{minipage}
  \begin{minipage}{0.48\textwidth}
    \ottdefnSINGLabeled
    \caption{Singular Constraint Entry}
    \label{fig:constraint-entry-singularity}
  \end{minipage}
\end{figure}
